---
title: "HW2"
author: "Kat Lopez"
date: "9/13/2017"
output: html_document
---

Data & Libraries
```{r}
library("dplyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("tidyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("plyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("tidyverse", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("lme4", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("broom", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("merTools", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")

```

Prepping data
```{r}
PDS <- read.csv("~/Dropbox/3rd Fall/Longitudinal data analysis/PDS_II_LI_reduced_09_01_17.csv") # wide format
View(PDS)


## converts from wide to long
PDS <- rename(PDS, c("Subid_fMRI"= "ID"))
PDS <- rename(PDS, c("CON_CON_WAVE1"= "CONCON_WAVE1"))
PDS <- rename(PDS, c("CON_CON_WAVE2"= "CONCON_WAVE2"))
PDS <- rename(PDS, c("CON_CON_WAVE3"= "CONCON_WAVE3"))
PDS <- rename(PDS, c("CON_FPN_WAVE1"= "CONFPN_WAVE1"))
PDS <- rename(PDS, c("CON_FPN_WAVE2"= "CONFPN_WAVE2"))
PDS <- rename(PDS, c("CON_FPN_WAVE3"= "CONFPN_WAVE3"))
View(PDS)
wide_to_long <- PDS %>%
gather(-ID, key = "Timepoint", value = "Value") %>%
separate(Timepoint, into = c("var", "omit", "wave"), sep = c(6, 11)) %>%
select(-omit) %>%
spread(key = var, value = Value)

wide_to_long <- rename(wide_to_long, c("Ageage" = "AgeatWave"))
View(wide_to_long)

## randomly created DOB and added to dataframe
Agenow=sample(seq(as.Date('2000/01/01'), as.Date('2012/01/01'), by='day'), 136)
DOBDOB_WAVE1=Agenow

#Centers age to be able to interpret 0

Agecen_WAVE1= PDS$Ageage_WAVE1- mean(PDS1$Ageage_WAVE1)
Agecen_WAVE2= PDS$Ageage_WAVE2- mean(PDS1$Ageage_WAVE2, na.rm=TRUE)
Agecen_WAVE3= PDS$Ageage_WAVE3- mean(PDS1$Ageage_WAVE3)
PDS1=cbind(PDS, Agecen_WAVE1, Agecen_WAVE2, Agecen_WAVE3, DOBDOB_WAVE1)
PDS1
View(PDS1)



wide_to_long_new <- PDS1 %>%
gather(-DOBDOB_WAVE1,-ID, key = "Timepoint", value = "Value") %>%
separate(Timepoint, into = c("var", "omit", "wave"), sep = c(6, 11)) %>%
spread(key = var, value = Value)
wide_to_long_new
View(wide_to_long_new)


```

### Homework Questions ###

# 1.Run linear models on all of your subjects (a basic regression). What is the average intercept, the average slope
```{r}
regressions_CONCON= lm (CONCON ~ Agecen, data = wide_to_long_new)
  summary(regressions_CONCON)
anova(regressions_CONCON)
# Average intercept is 0.297377 for connectivity between CON and CON networks of all individuals across time
# Average slope is 0.004034 of CON-CON 

regressions_CONFPN= lm (CONFPN ~ Agecen, data = wide_to_long_new)
  summary(regressions_CONFPN)
anova(regressions_CONFPN)
# Average intercept is -0.029210 for connectivity between CON and FPN networks of all individuals across time
# Average slope is -0.003149
```

#2. Now run a mlm/lmer model with only a random intercept. What is the ICC? What does residual variance look like compared to linear model? Create a graph to show this effect.

```{r}
## For CON-CON connectivity

Random_intercept_only= lmer(CONCON ~ 1 + (1 | ID), data = wide_to_long_new)
## predicts CON-CON connectivity from only the intercept but allows the intercept to vary across IDs
summary(Random_intercept_only)
## Fixed Effects (intercept) = 0.297052
## Residual Variance of MLM (variability of the intercept)= 0.005125
## Residual standard error for linear model = 0.08346 
### Overall, Residual Variance of the MLM is  smaller than the Residual standard error for linear model 

## calculating ICC= 0.269839
ICC= 0.001894/ (0.001894+ 0.005125)
ICC


# graphing mlm with only a random intercept. 
linearresidual= abs(broom::augment(regressions_CONCON)[,c(4)])
mean_linearresidual= mean(linearresidual)
SD_linearresidual= sd(linearresidual)

MLMresidual= abs(broom::augment(Random_intercept_only)[,c(4)])
mean_MLMresidual= mean(MLMresidual)
SD_MLMresidual= sd(MLMresidual)

linear_MLM= data.frame("Type"= c("linear", "MLM"), "mean"= c(mean_linearresidual, mean_MLMresidual), "SD"= c(SD_linearresidual, SD_MLMresidual))

ResidPlot= ggplot(linear_MLM, aes (Type, mean)) + geom_col() +geom_errorbar(aes(ymin= mean-SD, ymax= mean +SD), width= 0.2)
ResidPlot+ labs(y= "mean Residual", x= "Model Type")
#mean residuals appear similar, with MLM showing greater variance. 

## For CON-FPN connectivity 

model2= lmer(CONFPN ~ 1 + (1 | ID), data = wide_to_long_new)
summary(model2)
## Fixed Effects (intercept) = -0.02938087
## Residual Variance of MLM (Variance left over for each person)= 0.0019769
## Residual standard error for linear model = 0.05152 on 344 degrees of freedom
## For CON-FPN connectivity, relative to CONCON connectivity, Residual Variance of the MLM is much smaller than the  Residual standard error for linear model 

## calculating ICC= 0.2549559
ICC2= 0.0006765/ (0.0006765+ 0.0019769)
ICC2


# graphing mlm with only a random intercept. 
linearresidual2= abs(broom::augment(regressions_CONFNP)[,c(4)])
mean_linearresidual2= mean(linearresidual2)
SD_linearresidual2= sd(linearresidual2)

MLMresidual2= abs(broom::augment(model2)[,c(4)])
mean_MLMresidual2= mean(MLMresidual2)
SD_MLMresidual2= sd(MLMresidual2)

linear_MLM2= data.frame("Type"= c("linear", "MLM"), "mean"= c(mean_linearresidual2, mean_MLMresidual2), "SD"= c(SD_linearresidual2, SD_MLMresidual2))

ResidPlot2= ggplot(linear_MLM2, aes (Type, mean)) + geom_col() +geom_errorbar(aes(ymin= mean-SD, ymax= mean +SD), width= 0.2)
ResidPlot2 + labs(y= "mean Residual", x= "Model Type")
#mean residuals for the MLM is considerably smaller with wider variance than the linear model. 

```



# 3. Introduce a fixed slope term. What is the difference in terms of the fixed effects estimates between this estimate and the previous? Of the residual standard error? Create a graph to show both fixed effects estimates and the CIs around them.
```{r}
## FOR CONCON
fixed_slope1= lmer(CONCON ~ 1 + Agecen + (1 | ID), data = wide_to_long_new)
summary(fixed_slope1)

# fixed effect Intercept = 0.297126
# fixed effect Slope = 0.003605
# Residual error= 0.005131

# Both the fixed effect intercept and the residual standard errors increased from the previous model; fixed effect from 0.297052 to 0.297126 and the standard error from 0.005125 to 0.005131

confint(fixed_slope1, level = .95)
CI_CONCON= data.frame(confint(fixed_slope1, level = .95) [3:4,])
CI_CONCON
CI_CONCON$FE= fixef(fixed_slope1)
CI_CONCON_tidy=tidy(fixed_slope1)
CI_CONCON_tidy
dataframe1= data.frame("FixedEffect" = c ("intercept", "slope"),"Mean"= c(CI_CONCON[1,3], CI_CONCON[2,3]), "LowerCI"= c(CI_CONCON[1,1], CI_CONCON[2,1]), "UpperCI"= c(CI_CONCON[1,2], CI_CONCON[2,2]))
CI_graph= ggplot(dataframe1, aes(FixedEffect, Mean))+ geom_point() + geom_errorbar(aes(ymin= LowerCI, ymax= UpperCI), width=0.5)
CI_graph



## FOR CONFPN
fixed_slope2= lmer(CONFPN ~ 1 + Agecen + (1 | ID), data = wide_to_long_new)
summary(fixed_slope2)
# fixed effect Intercept = -0.029438
# fixed effect Slope = -0.002792
# Residual error =  0.0019780

# Fixed effect intercept decreased from -0.029381 in the previous model to -0.029438 in the current model. Residual variance also increased slightly from 0.0019769 to 0.0019780


confint(fixed_slope2, level = .95)
CI_CONFPN= data.frame(confint(fixed_slope2, level = .95) [3:4,])
CI_CONFPN
CI_CONFPN$FE= fixef(fixed_slope2)
CI_CONFPN_tidy=tidy(fixed_slope2)
CI_CONFPN_tidy
dataframe2= data.frame("FixedEffect" = c ("intercept", "slope"),"Mean"= c(CI_CONFPN[1,3], CI_CONFPN[2,3]), "LowerCI"= c(CI_CONFPN[1,1], CI_CONFPN[2,1]), "UpperCI"= c(CI_CONFPN[1,2], CI_CONFPN[2,2]))
CI_graph2= ggplot(dataframe2, aes(FixedEffect, Mean))+ geom_point() + geom_errorbar(aes(ymin= LowerCI, ymax= UpperCI), width=0.5)
CI_graph2

```


# 4. Run an additional model with a random slope. How does this change compare to the previous model? Should you keep the random slope or not?

```{r}
## FOR CONCON
randomslope1= lmer(CONCON~ 1 + Agecen+ (Agecen | ID), data = wide_to_long_new)
summary (randomslope1)

#previous model
# fixed effect Intercept = 0.297126
# fixed effect Slope = 0.003605
# Residual error= 0.005131


#model with random slope
# fixed effect Intercept = 0.296943
# fixed effect Slope = 0.003162
# Residual error = 0.005135
# random effects of slope= 5.511e-05

# Residual variance error stays nearly the same when introducing the random slope. It wouldnt make a difference to keep or remove the random slope.



## FOR CONFPN
randomslope2= lmer(CONFPN~ 1 + Agecen+ (Agecen | ID), data = wide_to_long_new)
summary (randomslope2)

# for previous model
# fixed effect Intercept = -0.029438
# fixed effect Slope = -0.002792
# Residual error =  0.0019780


# model with random slope
# fixed effect Intercept = -0.029639
# fixed effect Slope = -0.002333
# Residual= 0.0019683
# random slope of age= 50.0001159

# Residual variance decreased ( not by much)--I should retain the random slope


```

# 5. Interpret the correlation between the slope and the intercept.

A 0.163 correlation represents an average (and low) positive relationship between connectivity strength and age, such that connectivity becomes increasingly positive as a function of age. 



# 6. Create a density plot of the random effects from your final model.
```{r}

# FOR CONCON
# random intercept
CONCON_re.sim_intercept= REsim(Random_intercept_only)
head(CONCON_re.sim_intercept)
CONCON_density= CONCON_re.sim_intercept %>%
  filter(term == "(Intercept)")
ggplot(CONCON_density, aes(mean)) + geom_density()
# random slope
CONCON_re.sim_slope= REsim(randomslope1)
head(CONCON_re.sim_slope)
CONCON_density2= CONCON_re.sim_slope %>%
  filter(term == "(Intercept)")
ggplot(CONCON_density2, aes(mean)) + geom_density()

#FOR CONFPN

CONFPN_re.sim_intercept= REsim(model2)
head(CONFPN_re.sim_intercept)
CONFPN_density= CONFPN_re.sim_intercept %>%
  filter(term == "(Intercept)")
ggplot(CONFPN_density, aes(mean)) + geom_density()
# random slope
CONFPN_re.sim_slope= REsim(randomslope2)
head(CONFPN_re.sim_slope)
CONFPN_density2= CONFPN_re.sim_slope %>%
  filter(term == "(Intercept)")
ggplot(CONFPN_density2, aes(mean)) + geom_density()

```


# 7. Create a catepilar plot of the random effects. 
Is there any person that seems odd in terms of a large standard errors around intercept and slope estimates?
```{r}
# FOR CONCON
p1= plotREsim(CONCON_re.sim_intercept, labs=T)
p1

q1=plotREsim(CONCON_re.sim_slope, labs = T)
q1
## subjects 261 and 165 seems to have particularly large standard error around both the intercepts and slopes


#FOR CONFPN 
p2= plotREsim(CONFPN_re.sim_intercept, labs = T)
p2
q2=plotREsim(CONFPN_re.sim_slope, labs = T)
q2

## subjects 605 has larger standard error around both the intercept. 

```



# 8. Create a plot of the trajectory, along with a spaghetti plot of each person’s individual slope. Set the alpha level (transparency) on the individual slopes to make them easier to see.
```{r}
# For CONCON

Plot= ggplot(wide_to_long_new, aes(x= Agecen, y= CONCON, group=ID)) +geom_point() +stat_smooth(method = "lm", se= FALSE)
Plot2= Plot + stat_smooth(data= wide_to_long_new, aes (x=Agecen, y= CONCON, group=1, colour="red"), method="lm", se=FALSE)
Plot2

# For CONFPN

Plot1= ggplot(wide_to_long_new, aes(x= Agecen, y= CONFPN, group=ID)) +geom_point() +stat_smooth(method = "lm", se= FALSE)
Plot3= Plot1 + stat_smooth(data= wide_to_long_new, aes (x=Agecen, y= CONFPN, group=1, colour="red"), method="lm", se=FALSE)
Plot3

























