---
title: "SEM"
author: "Kat Lopez"
date: "October 31, 2017"
output:
  pdf_document: default
  html_document: default
---
```{r}
PDS <- read.csv("~/Documents/GitHub/1-descriptives-and-graphs-lopezkc/Connectivity Project/PDS.csv")
View(PDS)
library(lavaan)
library(semTools)
library("OpenMx", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library(semPlot)
library(lme4)
library(car)
library(tidyr)

```

1) Fit a measurement model to your constructs at one time point. Try out the different types of scaling discussed in class. What changes what stays the same?

```{r setup, include=FALSE}

Model1<- 'Executive_network=~ CONCON_WAVE1 + CONFPN_WAVE1 + CONVAN_WAVE1 +  CONDAN_WAVE1'

# Forces items to be on the same scale. Can be acheived one of two ways:

# Fixed factor; constrains variance of latent factor to 1
fixed<- cfa(Model1, std.lv=TRUE, data=PDS)
summary(fixed, fit.measures= TRUE, standardized= TRUE)
semPaths(fixed, what = "paths", whatLabels= "est", layout = "tree")

# Marker variable; fixes first factor loading (CONCON_WAVE1) to 1
marker<- cfa(Model1, data=PDS)
summary (marker, fit.measures=TRUE, standardized= TRUE)
semPaths(marker, what = "paths", whatLabels= "est", layout = "tree")

# All parameters stay the same with the exception of latent variable and variances estimates, which is expected as constraints differ between models 
```

2) What do the fit statistics say about your latent variable? Good/bad? Is your latent variable Just identified/saturated, under identified or over identified?

```{r setup, include=FALSE}
# Absolute indexes of model fit compare the fit of your model to a perfect fitting model (two below)
  # RMSEA= 0.034 # considered to be good fit
  # SRMR= 0.041 # considered to be good fit

# two below compare the fit of your model to the fit of the null model
  # CFI= 0.982 # considered to be good fit
  # TLI= 0.947 # considered to be good fit

# df= knowns-knowns; DF=2. This would  suggest that the latent variable is overidentified and thus we are able to assess the fit of the model and estimate the unknown parameters. However, p value of chi square is not significant (p=0.329)


```
3) Fit a longitudinal CFA model where you a) first correlate your latent factors across time and then b) a second model that predicts later times by a prevous time (ie auto regressive; t1 -> t2 -> t3). What are your conclusions? How does one differ from the other?

```{r setup, include=FALSE}
Full.mod<- '
Executive_network =~ CONCON_WAVE1 + CONFPN_WAVE1 + CONVAN_WAVE1 +  CONDAN_WAVE1
Executive_network2 =~ CONCON_WAVE2 + CONFPN_WAVE2 + CONVAN_WAVE2 +  CONDAN_WAVE2
Executive_network3 =~ CONCON_WAVE3 + CONFPN_WAVE3 + CONVAN_WAVE3 +  CONDAN_WAVE3
'

test<- cfa(Full.mod, data=PDS, std.lv=TRUE) 
summary (test, fit.measures=TRUE, standardized= TRUE)
semPaths(long_model, what = "paths", whatLabels= "est", layout = "tree")
# am getting a warning signal ' model has not converged'

aut0regressive<- 'Executive_network =~ CONCON_WAVE1 + CONFPN_WAVE1 + CONVAN_WAVE1 +  CONDAN_WAVE1
Executive_network2 =~ CONCON_WAVE2 + CONFPN_WAVE2 + CONVAN_WAVE2 +  CONDAN_WAVE2
Executive_network3 =~ CONCON_WAVE3 + CONFPN_WAVE3 + CONVAN_WAVE3 +  CONDAN_WAVE3
Executive_network3 ~ Executive_network2
Executive_network2 ~ Executive_network
'

autoreg<- cfa(aut0regressive, std.lv=TRUE, data=PDS) 
summary (autoreg, fit.measures=TRUE, standardized= TRUE)
semPaths(aut0regressive, what = "paths", whatLabels= "est", layout = "tree")

### am getting a warning signal ' model has not converged'###


```

4) Fit a longitdinal growth model in SEM and in HLM. Compare and contrast the differences.

Intercept between SEM and HLM were largely the same, slopes were similar as well. Introducing a random slope in HLM did not change these figures. 

```{r setup, include=FALSE}
  #prep data for HLM
    PDS <- read.csv("~/Dropbox/3rd Fall/Longitudinal data analysis/PDS_II_LI_reduced_09_01_17.csv")
    
    ## renaming variables
    PDS <- rename(PDS, c("Subid_fMRI"= "ID"))
    PDS <- rename(PDS, c("CON_CON_WAVE1"= "CONCON_WAVE1"))
    PDS <- rename(PDS, c("CON_CON_WAVE2"= "CONCON_WAVE2"))
    PDS <- rename(PDS, c("CON_CON_WAVE3"= "CONCON_WAVE3"))
    PDS <- rename(PDS, c("CON_FPN_WAVE1"= "CONFPN_WAVE1"))
    PDS <- rename(PDS, c("CON_FPN_WAVE2"= "CONFPN_WAVE2"))
    PDS <- rename(PDS, c("CON_FPN_WAVE3"= "CONFPN_WAVE3"))

    ## randomly created DOB and added to dataframe
    Agenow=sample(seq(as.Date('2000/01/01'), as.Date('2012/01/01'), by='day'), 136)
    DOBDOB_WAVE1=Agenow
    
    #Centers age to be able to interpret 0
    PDS1=PDS
    Agecen_WAVE1= PDS$Ageage_WAVE1- mean(PDS1$Ageage_WAVE1)
    Agecen_WAVE2= PDS$Ageage_WAVE2- mean(PDS1$Ageage_WAVE2, na.rm=TRUE)
    Agecen_WAVE3= PDS$Ageage_WAVE3- mean(PDS1$Ageage_WAVE3)
    PDS1=cbind(PDS, Agecen_WAVE1, Agecen_WAVE2, Agecen_WAVE3, DOBDOB_WAVE1)
    ## converting to long format
    wide_to_long_new <- PDS1 %>%
    gather(-DOBDOB_WAVE1,-ID,-Sex, key = "Timepoint", value = "Value") %>%
    separate(Timepoint, into = c("var", "omit", "wave"), sep = c(6, 11)) %>%
    spread(key = var, value = Value)

#HLM growth model random intercept; intercept=0.297126, slope= 0.003605
HLM<- lmer(CONCON ~ Agecen+ (1|ID), data=wide_to_long_new)
summary(HLM)

# HML growth model with random slope; intecept= 0.296943, slope= 0.003162
randomslope1= lmer(CONCON~ 1 + Agecen+ (Agecen | ID), data = wide_to_long_new)
summary (randomslope1)

# SEM growth model with 3 timepoints intercept=0.291 and slope=0.008 with fixed coefficients  
growth.model<- ' i=~ 1*CONCON_WAVE1 + 1*CONCON_WAVE2 + 1*CONCON_WAVE3
s=~ 0*CONCON_WAVE1 + 1*CONCON_WAVE2 + 2*CONCON_WAVE3'
fit<- growth(growth.model, data = PDS)
summary(fit)
semPaths(fit, what = "paths", whatLabels= "est", layout = "tree")

# Intercepts are nearly identical, HLM model has slightly smaller slope.
```

5) Constrain the residual variances to be equal. Does this change the fit of your model? Intercept= 0.291, Slope= 0.006, not much change.

```{r setup, include=FALSE}
constrain.var<- ' i=~ 1*CONCON_WAVE1 + 1*CONCON_WAVE2 + 1*CONCON_WAVE3
s=~ 0*CONCON_WAVE1 + 1*CONCON_WAVE2 + 2*CONCON_WAVE3
CONCON_WAVE1~~ a*CONCON_WAVE1
CONCON_WAVE2~~ a*CONCON_WAVE2
CONCON_WAVE3~~ a*CONCON_WAVE3
'
constrain.var1<- growth(constrain.var, missing= "ML", data = PDS)
## error: lavaan WARNING: some estimated lv variances are negative ##
summary(constrain.var1)
semPaths(constrain.var1, what = "paths", whatLabels= "est", layout = "tree")
```

6) Contrain your slope to be fixed, not random. How does this change your model?

```{r setup, include=FALSE}
#HML growth model with fixed slope; intercept=  0.297126, slope= 0.003605

fixed_slope1= lmer(CONCON ~ 1 + Agecen + (1 | ID), data = wide_to_long_new)
summary(fixed_slope1)
```

7) Change the time metric in your SEM growth model. How does that change your estimates? Does it change your fit statistics?


```{r setup, include=FALSE}
```

8) Try a different type of estimation (see lavaan tutorial for details). How does that change your model?

```{r setup, include=FALSE}
```
